### 2018-12-10 Revision/Response/response.tex

%We thank the reviewers for their assessment of our manuscript. Please find
%below point-by-point replies (in black) to your comments and questions
%(reprinted in blue). Where possible, key modifications have been highlighted
%in blue in the revised manuscript. We sincerely hope that with these
%modifications, the paper is found suitable for publication in the
%{\it Journal of Scientific Computing}.

We are very thankful to the reviewers for recognizing potential in this effort
and for their constructive feedback on the manuscript. A point-by-point
response to the comments has been provided below. Corresponding modifications
have been highlighted (in blue) in the revised manuscript. Major changes in the
revised manuscript are summarized as follows:

\begin{itemize}

\item The complexity of the problem has been increased substantially in two ways: First, 
the non-linearity of the system response has been significantly increased by increasing
the perturbations associated with the uncertain rate-controlling parameters: pre-exponent
and the activation energy. Second, the dimensionality of the problem for the higher
dimensional case has been increased from 33 to 36.
Specifically, in addition to the rate parameters, the impact of
perturbing the initial conditions: pressure, temperature, and stoichiometry pertaining to the
H$_2$/O$_2$ reaction mechanism on the ignition delay is investigated. Such analyses using the
proposed framework would provide useful insights to combustion scientists interested in
kinetics applications. Moreover, it illustrates the applicability and robustness of the 
proposed framework to complex kinetics applications.

@@

As mentioned in Section 6
(Summary and Discussion), the proposed framework is agnostic to the choice of the system and could be extended
to other kinetics applications as long as the system response is continuously differentiable in the domain of
the inputs and is therefore applicable to a wide variety of applications.

\item \referee{Related to 1) above, it is of paramount importance that the authors consider, at least, 
methane, for which I expect less of a spectral gap in the matrix C. Ideally, more complex fuels with 
more nuanced responses.
In other words, if the method works well for a simple case, what is its value to the community?
The authors must try this out on a much more complex hydrocarbon/oxygen system.}

A more complex problem as suggested by the reviewer would manifest itself into an increased number of
uncertain parameters and/or an increased amount of non-linearity in the system response. Therefore, while
we did not implement the framework to a new system, we have significantly increased the non-linearity of the
response by increasing the degree of uncertainty in the rate parameters. Specifically, instead of
the pre-exponent, $A_i$, we perturb $\log(A_i)$ as suggested by the other reviewer. Moreover, 
the magnitude of the perturbation has been increased from 1$\%$ to $2\%$ in the
higher-dimensional case.
In addition to the rate parameters,
we perturb the initial conditions to investigate their impact on the response individually as well as due to
their interaction with the rate parameters. Hence, we have also increased the dimensionality of the problem
for the higher dimensional case from 33 to 36. More importantly, the impact of initial conditions on the
variability in the system response could provide useful insights to combustion scientists interested in 
reaction kinetics for a range of applications.

### 2018-12-10 Revision/Response/response.tex

We agree; we have removed 'statistical' from that part.

\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is
just the negative log-PDF, I am not sure if it is necessary to bring up a
Boltzmann distribution language here.  Also, it could be confused with
Maxwell-Boltzmann distribution, which is a very specific PDF form.}

Thanks for
this comment. We followed the theory in ref [35] in the manuscript, where the
authors derive Poincare type inequalities for functions of iid random inputs
distributed according to Boltzmann probability laws.  We point out that the
uniform distribution does not belong to this class.  The required Poincare type
inequality for the case of uniform distribution was derived in an earlier paper
of Sobol' and Kucherenko (ref [34]).  That was the reason for assuming the
distribution laws of the random inputs are either \emph{uniform} or belong to
the Boltzmann class, in section 3. An important class of Boltzmann measures are
the so called log-concave distributions such as Normal, Exponential, Beta, and
Gamma distributions. These points have been clarified in the revised
manuscript.

Additionally, we share the reviewer's concern about the potential confusions
some readers might face when seeing the mention of Boltzmann. We have removed
the mention of Boltzmann measures from the introduction, and now only refer to
Boltzmann measures in Section 3, where the theoretical results are discussed.

### 2018-12-10 Revision/Paper/app.tex
..

For the high-dimensional case, we aim to investigate the impact of 
uncertainty in the following problem parameters on 
the ignition delay associated with the H$_2$/O$_2$ reaction:
(i) pre-exponents ($A_i$'s); (ii) the activation energies
($E_{a,i}$'s); and (iii) \rebut{the initial pressure~($P_0$),
temperature~($T_0$), and stoichiometry~($\Phi_0$)}. 
%
\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except $\mathcal{R}_6$
-- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to zero nominal values for $E_a$),
and the initial conditions were perturbed by 2$\%$ about their nominal values.
Note that the magnitude of the perturbation was selected such that the
ignition delay assumes a physically meaningful value in the input domain. 
The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
$\Phi_0$ were considered to be 1.0~atm, 900~K, and 2.0 respectively.}

>>>>

For the high-dimensional case, we aim to investigate the impact of 
uncertainty in the following problem parameters on 
the ignition delay associated with the H$_2$/O$_2$ reaction:
(i) pre-exponents ($A_i$'s); (ii) the activation energies
($E_{a,i}$'s); and (iii) \rebut{the initial pressure~($P_0$),
temperature~($T_0$), and stoichiometry~($\Phi_0$)}. 
%
\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except $\mathcal{R}_6$
-- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to zero nominal values for $E_a$),
and the initial conditions were considered to be uniformly distributed, and
perturbed by 2$\%$ about their nominal values.
Note that the magnitude of the perturbation was selected such that the
ignition delay assumes a physically meaningful value in the input domain. 
The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
$\Phi_0$ were considered to be 1.0~atm, 900~K, and 2.0 respectively.}

### 2018-12-10 Revision/Paper/method.tex
...
..

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to analyze 
the H$_2$/O$_2$ reaction kinetics
problem whereby the \rebut{logarithm of the} 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
$A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Two approaches are explored for estimating the gradient of ignition delay with
respect to \rebut{$\log(A_i)$}: 
a perturbation  approach that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression  approach 
that involves a linear regression fit to the available set of
model evaluations in order to approximate the gradient. 
The active subspace is computed in an iterative manner to avoid
unnecessary model evaluations once converged is established.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp}~(right).
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
%\caption{\rebut{An illustrative comparison of the SSPs generated using the 
%perturbation and the regression strategies for computing the active subspace.}}
%\label{fig:comp_ssp}
%\end{center}
%\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}

>>>>

In this section, we outline the methodology for computing the
active subspace in an efficient manner. The proposed framework is employed to analyze 
a 19-dimensional H$_2$/O$_2$ reaction kinetics
problem whereby the \rebut{logarithm of the} 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
$A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Two approaches are explored for estimating the gradient of ignition delay with
respect to \rebut{$\log(A_i)$}: 
a perturbation  approach that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression  approach 
that involves a linear regression fit to the available set of
model evaluations in order to approximate the gradient. 
The active subspace is computed in an iterative manner to avoid
unnecessary model evaluations once converged is established.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp}~(right).
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
%\caption{\rebut{An illustrative comparison of the SSPs generated using the 
%perturbation and the regression strategies for computing the active subspace.}}
%\label{fig:comp_ssp}
%\end{center}
%\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on
$N$~=~45 samples which requires $M$~=~900 model runs. For the same amount of
computational effort, we can afford $N$~=~900 samples when using regression.
Hence, the SSP from regression is based on 900 points:
($\hat{\mat{W}}_1^\top\bm{\xi}_j$,~$G(\hat{\mat{W}}_1^\top\bm{\xi}_j)$),
$j = 1, \ldots, 900$. On the
other hand, the SSP from perturbation is plotted using only 45 points as
mentioned earlier.  Nevertheless, the illustrative comparison clearly indicates
that the two SSPs are in excellent agreement. Moreover, it is interesting to
note that the response in ignition delay based on the considered probability
distributions for $\log(A_i)$ although non-linear, can be approximated by a
1-dimensional active subspace.}
